---
title: Regression Tables w/ Many Subsets of Data in Long Form
author: R package build
date: '2021-11-08'
slug: regression-tables-w-many-subsets-of-data
categories:
  - cleaning
  - R
  - modelsummary
  - kable
  - kableExtra
tags:
  - regression
  - tips
  - wrangling
  - presentation
subtitle: ''
summary: 'Make a unique table that is completely automated!'
authors: []
lastmod: '2021-11-08T20:12:51-08:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
editor_options: 
  chunk_output_type: console
---
*Please turn the website to "light mode" to see the tables clearly! *

**Update (10/3/2022):** The steps in this tutorial can be easily replicated with one simple function using my new package called [Panelsummary](https://michaeltopper1.github.io/panelsummary/index.html). 


Oftentimes, it is preferable to create a regression table that has many subsets of your data set. For instance, in my own work on fraternities and crime, I find it necessary to show three sets of results: effects on in the full sample, effects on the weekends, and effects on the weekdays. Moreover, I believe a great way to present this data is with a *long* table rather than a *wide* table. In particular, this means having a table with grouped rows—one grouping for each subset of data. Usually, these tables can only be created with manual editing since no out-out-box function in R can handle such a task at the moment. However, the `modelsummary` package can actually achieve this with some small data manipulation with `dplyr`. For reference, below is a regression table I'll demonstrate how to make:

```{r, echo=F, message = F, warning = F}
library(fixest)
library(modelsummary)
library(tidyverse)
library(kableExtra)
cars <- mtcars %>% 
  mutate(mpg_levels = case_when(
    between(mpg, 0, 20) ~ "Low MPG",
    between(mpg, 20, 25) ~ "Mid MPG",
    between(mpg, 26, 50) ~ "High MPG"
  )) 

low_mpg <- cars %>% 
  filter(mpg_levels == "Low MPG") %>% 
  feols(mpg ~ hp |gear , data = .)
low_mpg_2 <- cars %>% 
  filter(mpg_levels == "Low MPG") %>% 
  feols(mpg ~ hp |gear + am , data = .)

mid_mpg <- cars %>% 
  filter(mpg_levels == "Mid MPG") %>% 
  feols(mpg ~ hp |gear , data = .)
mid_mpg_2 <- cars %>% 
  filter(mpg_levels == "Mid MPG") %>% 
  feols(mpg ~ hp |gear + am, data = .)

high_mpg <- cars %>% 
  filter(mpg_levels == "High MPG") %>% 
  feols(mpg ~ hp |gear , data = .)
high_mpg_2 <- cars %>% 
  filter(mpg_levels == "High MPG") %>% 
  feols(mpg ~ hp |gear + am , data = .)

low <- list(low_mpg, low_mpg_2)
mid <- list(mid_mpg, mid_mpg_2)
high <- list(high_mpg, high_mpg_2)


gm_first <- tribble(~raw, ~clean, ~fmt,
                      "nobs", "Num.Obs", ~fmt)

gm_last <- tribble(~raw, ~clean, ~fmt,
                      "nobs", "Num.Obs", ~fmt,
                   "FE: gear", "FE: Gear", ~fmt,
                   "FE: am", "FE: AM", ~fmt)
low_table <- modelsummary(low, stars = T, 
             coef_map = c("hp" = "Horse Power"), output = 
               "data.frame",
             gof_map = gm_first)  %>% 
  mutate(term = ifelse(statistic == "modelsummary_tmp2", "", term)) %>%
  select(matches("term|^model"))

mid_table <- modelsummary(mid, stars = T, 
             coef_map = c("hp" = "Horse Power"), output = 
               "data.frame",
             gof_map = gm_first)  %>% 
  mutate(term = ifelse(statistic == "modelsummary_tmp2", "", term)) %>%
  select(matches("term|^model"))

high_table <- modelsummary(high, stars = T, 
             coef_map = c("hp" = "Horse Power"), output = 
               "data.frame",
             gof_map = gm_last)  %>% 
  mutate(term = ifelse(statistic == "modelsummary_tmp2", "", term)) %>%
  select(matches("term|^model"))

table <- bind_rows(low_table, mid_table, high_table)


final_table <- table %>% 
  kbl(col.names = c(" ", "(1)", "(2)"), booktabs = T,
      caption  = "Effect of Horse Power on MPG") %>% 
  kable_classic(full_width = F, html_font = "Cambria") %>% 
  pack_rows("Low MPG Cars", 1, 3, bold = F, italic = T) %>% 
  pack_rows("Mid MPG Cars", 4, 6, bold = F, italic = T) %>% 
  pack_rows("High MPG Cars", 7,9, bold = F, italic = T) %>% 
  pack_rows("Fixed Effects", 10, 11, bold = F, italic = T) %>% 
  add_header_above(c(" " = 1, "Outcome: MPG" = 2)) %>% 
  footnote(list("Low MPG cars are below 20 miles-per-gallon.", 
                "Mid MPG cars are between 20 and 25 miles-per-gallon.",
                "High MPG cars are above 25 miles-per-gallon.")) 
```

```{r, echo = F}
final_table
```

Note that this table will look even better when rendered to a PDF. 

Before we begin, load in the necessary packages:

```{r}
## loading in necessary packages
library(fixest)
library(modelsummary)
library(tidyverse)
library(kableExtra)
```


## Step 1: The Models

I am going to estimate two separate models using the `mtcars` dataset that comes pre-loaded in R. The first model will be a linear regression of `mpg` on `hp` with `gear` fixed effects. The second model will be a linear regression of `mpg` on `hp` but with `gear` AND `am` fixed effects. I'll split up the data into three subsets that I want to estimate these models on: cars with a low, mid, and high mpg. These models are not supposed to be insightful in any way, just a way to demonstrate the table-making process, so bear with me.

First, let's create a new column that will split our data into our desired subsets:

```{r}
## creating a new variable
cars <- mtcars %>% 
  mutate(mpg_levels = case_when(
    between(mpg, 0, 20) ~ "Low MPG",
    between(mpg, 20, 25) ~ "Mid MPG",
    between(mpg, 26, 50) ~ "High MPG"
  )) 

```

Now let's estimate our regressions. Use your favorite regression package, although I'll use `fixest` here:
```{r}

## Estimating both models on low mpg cars
low_mpg <- cars %>% 
  filter(mpg_levels == "Low MPG") %>% 
  feols(mpg ~ hp |gear , data = .)
low_mpg_2 <- cars %>% 
  filter(mpg_levels == "Low MPG") %>% 
  feols(mpg ~ hp |gear + am , data = .)

## Estimating both models on mid mpg cars
mid_mpg <- cars %>% 
  filter(mpg_levels == "Mid MPG") %>% 
  feols(mpg ~ hp |gear , data = .)
mid_mpg_2 <- cars %>% 
  filter(mpg_levels == "Mid MPG") %>% 
  feols(mpg ~ hp |gear + am, data = .)

## Estimating both models on high mpg cars
high_mpg <- cars %>% 
  filter(mpg_levels == "High MPG") %>% 
  feols(mpg ~ hp |gear , data = .)
high_mpg_2 <- cars %>% 
  filter(mpg_levels == "High MPG") %>% 
  feols(mpg ~ hp |gear + am , data = .)

```

Now, I'm going to save each of these groups of models into a list:

```{r}
## saving the low mpg models into a list
low <- list(low_mpg, low_mpg_2)

## saving the mid mpg models into a list
mid <- list(mid_mpg, mid_mpg_2)

## saving the high mpg models into a list
high <- list(high_mpg, high_mpg_2)
```

## Step 2: Pass these into `modelsummary`

The next step is to pass these into the `modelsummary` function, but with a slight twist. In particular, we are going to use the `output = "data.frame"` argument to create a dataframe of our models. But first, I want to note that I am going to omit all of the goodness-of-fit statistics that `modelsummary` spits out except the number of observations. To do this, I am going to use the `gof_map` argument. Take a look at the [modelsummary website](https://vincentarelbundock.github.io/modelsummary/articles/modelsummary.html) for how to use this, or you can check out this [stackoverflow conversation](https://stackoverflow.com/questions/69420683/is-there-a-way-to-change-the-capitalization-in-modelsummary-for-fixed-effects-wi) for a minimal reproducible example since I will not go into details here. The main takeaway for these next few lines of code is this: I am creating "cleaned" versions of the goodness-of-fit statistics that I want- the first is for the first two lists of models (`low` and `mid`), while the second is for the last list of models (`high`). I do this because I want the table to only have fixed effects at the bottom of the table, yet I want to have the number of observations within the table.

```{r}
## creating cleaned goodness-of-fit mappings for the low and mid models
gm_first <- tribble(~raw, ~clean, ~fmt,
                      "nobs", "Num.Obs", ~fmt)

## creating cleaned goodness-of-fit mappings for the high models
gm_last <- tribble(~raw, ~clean, ~fmt,
                      "nobs", "Num.Obs", ~fmt,
                   "FE: gear", "FE: Gear", ~fmt,
                   "FE: am", "FE: AM", ~fmt)
```

Next, I will pass in the `low`, `mid`, and `high` lists into modelsummary, and save the output as a dataframe.

```{r}
## saving low models as a data frame - doing some additional presentation cleaning
low_table <- modelsummary(low, stars = T, 
             coef_map = c("hp" = "Horse Power"), output = 
               "data.frame",
             gof_map = gm_first)  %>% 
  mutate(term = ifelse(statistic == "modelsummary_tmp2", "", term)) %>%
  select(matches("term|^model"))

## saving mid models as a data frame - doing some additional presentation cleaning
mid_table <- modelsummary(mid, stars = T, 
             coef_map = c("hp" = "Horse Power"), output = 
               "data.frame",
             gof_map = gm_first)  %>% 
  mutate(term = ifelse(statistic == "modelsummary_tmp2", "", term)) %>%
  select(matches("term|^model"))

## saving high models as a data frame - doing some additional presentation cleaning
## BE AWARE I PASSED IN GM_SECOND to the GOF_MAP arugment here!!!!
high_table <- modelsummary(high, stars = T, 
             coef_map = c("hp" = "Horse Power"), output = 
               "data.frame",
             gof_map = gm_last)  %>% 
  mutate(term = ifelse(statistic == "modelsummary_tmp2", "", term)) %>%
  select(matches("term|^model"))
```

This created 3 separate dataframes with the information we want. Note that I used the `gm_last` for the final `high_table` dataframe. Now, let's append these dataframes together. WARNING: Please be advised that the order *does** matter here since we want the fixed effects to be displayed last!!

```{r}
## binding together the rows - order matters!!
table <- bind_rows(low_table, mid_table, high_table)

```

## Step 3: Use `kableExtra` to customize

Now that you have your final dataframe `table`, you can pass this into `kableExtra::kbl` and begin customizing the table to your liking. I won't go into details, but the [kableExtra vignette](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html) is very detailed and excellent. 

```{r}
table %>% 
  kbl(col.names = c(" ", "(1)", "(2)"), booktabs = T,
      caption  = "Effect of Horse Power on MPG") %>%  ## cleaning columnnames and adding caption
  kable_classic(full_width = F, html_font = "Cambria") %>%  ## creating a latex-like output style
  pack_rows("Low MPG Cars", 1, 3, bold = F, italic = T) %>% ## grouping rows
  pack_rows("Mid MPG Cars", 4, 6, bold = F, italic = T) %>% ## grouping rows
  pack_rows("High MPG Cars", 7,9, bold = F, italic = T) %>% ## grouping rows
  pack_rows("Fixed Effects", 10, 11, bold = F, italic = T) %>%  ## grouping rows
  add_header_above(c(" " = 1, "Outcome: MPG" = 2)) %>%  ## adding header
  footnote(list("Low MPG cars are below 20 miles-per-gallon.", 
                "Mid MPG cars are between 20 and 25 miles-per-gallon.",
                "High MPG cars are above 25 miles-per-gallon."))  ## adding footnotes
```

And there you have it! If you know you are going to be using a similar-style table for a project, I would recommend writing a function that can handle step 2 automatically since the coding can get a little tedious. Feel free to email me with any questions at michaeltopper@ucsb.edu!
